---
title: "September14,2017"
author: "Garrett"
date: "9/14/2017"
output: html_document
---

```{r "cleanup", echo = FALSE, eval=FALSE}
# not run in knitted doc
rm(list = ls())
.rs.restartR()
```


```{r "load libraries", echo = FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(plotly)
library(stringr)
library(knitr)
library(readxl)
library(tidyr)
library(purrr)
#library(xlsx)
opts_chunk$set(dev=c("png", "pdf"), dev.args=list(pdf = list(encoding="WinAnsi", useDingbats=FALSE)),
  fig.keep="all", fig.path = file.path("plots", "smoky_hollows_SRM")) ### CHANGE PLOT SUFFIX ###
```

#Read and organize data
```{r}
read_srm_export <- function(filename, columns = c("peak_name", "RT.min", "basepeak", "area.cpm", "height.cts", "quantitation")) {
  filename %>% 
    # read excel files
    read_excel(sheet = "Integration", skip = 42, 
               col_names = columns, col_types = rep("text", length(columns))) %>% 
    as_data_frame() %>%
    # remove empty rows
    filter(!is.na(peak_name), peak_name != "n.a.") %>% 
    # convert the relevant numeric columns into numbers
    mutate_at(vars(RT.min, area.cpm, height.cts), as.numeric) %>% 
    # remove useless columns
    select(-basepeak, -quantitation) %>% 
    # add filename info
    mutate(file_id = gsub("\\.xls", "", basename(filename))) %>% 
    select(file_id, everything())
}

# get data
all_data <- 
  # find all excel files
  list.files("data_reports", recursive = TRUE, full.names = TRUE, pattern = "\\.xls$") %>% 
  # send them to the read method
  lapply(read_srm_export) %>% 
  # combine the data set
  bind_rows() %>% 
  # pull out sample information
  #mutate(sample_id = str_match(all_data$file_id, "TSQ\\d+_GB_(.*)$") %>% { .[,2] }) %>% 
  # get n replicates
  group_by(file_id)
  #mutate(n_replicates = length(unique(file_id)))

```

#File names for metadata file
```{r, echo=FALSE, eval=FALSE}
# file names
all_data$file_id %>% unique() %>%  paste(collapse = "\n") %>% message()
```

#Plot to check the spread
##by sequence
```{r}
all_data %>% 
  ggplot() + 
  aes(x = peak_name, y = area.cpm, color = file_id) +
  geom_point(size = 3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
##by sample
```{r}
all_data %>% 
  ggplot() +
  aes(x = file_id, y = area.cpm, color = file_id) + 
  geom_point(size = 3) +
  facet_wrap(~peak_name, scales = "free_y") +
    theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "none")
```

# Calculation peak amounts and rock concentrations

```{r}
depth_and_rock_info <- read_excel(file.path("metadata", "aromaticSRM_20180327.xlsx")) %>% 
  rename(rock.g = `rock .g`, tle = `TLE.mg`, maltene = `maltenes.mg`)%>% 
  filter(!is.na(file_id))
kable(depth_and_rock_info)

data_by_depth <- 
  all_data %>%
  left_join(depth_and_rock_info, by = "file_id") %>% 
  group_by(file_id) %>% 
  filter(`process` == "yes")  %>%
  mutate(
    n_peaks = n(),
    n_standards = sum(peak_name == "d14-pTerph"),
    ref_area.cpm = area.cpm[peak_name == "d14-pTerph"],
    ref_amount_added.ug = 5, #check standard amount added
    amount.ug = area.cpm/ref_area.cpm * ref_amount_added.ug,
    
     ##Normalize by what you want!
    conc_rock.ug_g = amount.ug / rock.g, 
    conc_tle.ug.g = amount.ug / tle,  
    conc_maltene.ug.g = amount.ug / maltene
    #total_area.cpm = sum(area.cpm[peak_name != "d14-pTerph"]),
    #area.percent = area.cpm / total_area.cpm * 100
  )%>% ungroup() %>% 
  arrange(file_id, peak_name)

View(data_by_depth)
```

# Calculate Recovery

### Linear regressions of the calibration curves

```{r}
standard <- read_excel(file.path("metadata", "D14 calibration.xlsx"))   ###read excel

###calibration curve
standard %>% 
  ggplot() +
  aes(x = Known.ng, y = Measured_area.counts, color = calibration) + 
  geom_smooth(method = "lm", alpha = 0.5) +
  geom_point() +
  theme_bw() +
  theme(legend.position = "none") 

calibrations <- 
  standard %>% 
  filter(!is.na(calibration)) %>% 
  nest(-calibration) %>% 
  mutate(
    fit = map(data, ~summary(lm(`Measured_area.counts`~ `Known.ng`, data = .x))),
    coefficients = map(fit, "coefficients"),
    intercept = map_dbl(coefficients, `[`, 1, 1),
    intercept_se = map_dbl(coefficients, `[`, 1, 2),
    slope = map_dbl(coefficients, `[`, 2, 1),
    slope_se = map_dbl(coefficients, `[`, 2, 2),
    r2 = map_dbl(fit, "r.squared")
  )

calibrations %>% select(-data, -fit, -coefficients) %>% knitr::kable(d = 3)
```

### Determine yield

These numbers are not useful for anything else.

```{r}
calib_data <-
  data_by_depth %>% 
  # temp
  mutate(calibration = "oct2017") %>% 
  left_join(calibrations, by = "calibration") %>% 
  mutate(
    total_volume.uL = 100,
    total_inject.uL = 1,
    ref_amount_inject_expected.ng = ref_amount_added.ug/total_volume.uL * total_inject.uL * 1000,
    ref_amount_inject_measured.ng = (ref_area.cpm - intercept)/slope,
    ref_amount_measured.ug = total_volume.uL/total_inject.uL * ref_amount_inject_measured.ng * 1/1000,
    yield = ref_amount_inject_measured.ng/ref_amount_inject_expected.ng
  )
  
```

#### check yields

```{r}
calib_data %>% 
  select(file_id, yield)  %>% 
  arrange(file_id)  %>% 
  unique() %>% 
  ggplot() + aes(file_id, y = 100*yield) +
  geom_point(size = 3) +
  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5))
```



#Combine compounds/make ratios (new rows w/o RTs etc, just concentration.rock column)

```{r}
# functions to make it easy to sum up peaks

sum_peaks <- function(df, filter_condition, new_peak_name) {
  filter_condition <- sprintf("(%s)", str_c(filter_condition, collapse = "|"))
  filter(df, str_detect(peak_name, filter_condition)) %>% 
    summarize(
      file_id = file_id[1],
      depth = depth[1],
      conc_tle.ug.g = sum(conc_tle.ug.g)
    ) %>% 
    mutate(peak_name = new_peak_name)
}

ratio_peaks <- function(df, filter_top, filter_bottom, new_peak_name) {
  filter_top <- sprintf("(%s)", str_c(filter_top, collapse = "|"))
  filter_bottom <- sprintf("(%s)", str_c(filter_bottom, collapse = "|"))
  filter(df, str_detect(peak_name, filter_top) | str_detect(peak_name, filter_bottom)) %>% 
    summarize(
      file_id = file_id[1],
      depth = depth[1],
      ratio = sum(conc_tle.ug.g[str_detect(peak_name, filter_top)]) / sum(conc_tle.ug.g[str_detect(peak_name, filter_bottom)])
    ) %>% 
    mutate(peak_name = new_peak_name)
}
```


```{r}
final_data <- data_by_depth %>% 
    group_by(file_id) %>% 
        do({
          bind_rows(., 
              sum_peaks(., "C15", "all C15"),
              sum_peaks(., "C16", "all C16"),
              sum_peaks(., "C17", "all C17"),
              sum_peaks(., "C18", "all C18"),
              sum_peaks(., "C19", "all C19"),
              sum_peaks(., "C20", "all C20"),
              sum_peaks(., "C21", "all C21"),
              sum_peaks(., "C24", "all C24"),
              sum_peaks(., "C26", "all C26"),
              sum_peaks(., "Aryl", "all_Aryl_Isop"),
              #sum_peaks(., "MP", "3ring_MP"),
              sum_peaks(., c("Acenapthene", "Flourene"), "2ring_all" ),
              #sum_peaks(., "Acenapthene", "2ring_1"),
              #sum_peaks(., "Flourene", "2ring_2"),
              sum_peaks(., c("Phenanthrene", "Flourantherene", "Retene", "MP"), "3ring_all" ),
              #sum_peaks(., "Phenanthrene", "3ring_1"),
              #sum_peaks(., "Flourantherene", "3ring_2"),
              #sum_peaks(., "Retene", "3ring_3"),
              sum_peaks(., c("Pyrene", "Benzo[a]anthracene", "Triphenylene", "Chrysene", "flouranthrene"), "4ring_all" ),
              #sum_peaks(., "Pyrene", "4ring_1"),
              #sum_peaks(., "Benzo[a]anthracene", "4ring_2"),
              #sum_peaks(., "Triphenylene", "4ring_3"),
              #sum_peaks(., "Chrysene", "4ring_4"),
              #sum_peaks(., "flouranthrene", "4ring_5s"),
              sum_peaks(., c("Benzo[e]pyrene", "Benzo[a]pyrene", "Perylene", "Ideno", "Dibenz"), "5ring_all" ),
              #sum_peaks(., "Benzo[e]pyrene", "5ring_1"),
              #sum_peaks(., "Benzo[a]pyrene", "5ring_2"),
              #sum_peaks(., "Perylene", "5ring_3"),
              #sum_peaks(., "Ideno[c,e]", "5ring_4"),
              #sum_peaks(., "Dibenz[a,h]", "5ring_5"),
              sum_peaks(., c("Benzo[ghi]", "Coronene"), "6ring_all" ), 
              sum_peaks(., c("Acenapthene", "Flourene", "Phenanthrene", "Flourantherene", "Retene", "Pyrene", "Benzo[a]anthracene", "Triphenylene", "Chrysene", "flouranthrene", "Benzo[e]pyrene", "Benzo[a]pyrene", "Perylene", "Ideno[c,e]", "Dibenz[a,h]", "Benzo[ghi]", "Coronene"), "all_PAH")
              #sum_peaks(., "Benzo[ghi]", "6ring_1"),
              #sum_peaks(., "Coronene", "6ring_2")
 ) }) %>% ungroup() 
  
final_data


```



#### Against depth

```{r "overview_against_depth", fig.width = 10, fig.height = 8}
final_data %>% 
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth, color = file_id) +
  geom_point() +
  facet_wrap(~peak_name, scales = "free") +
  scale_y_reverse() 

```



##By compound
###D14
```{r "D14_against_depth", fig.width = 10, fig.height = 8}
subset(final_data, peak_name=='d14-pTerph') %>% 
    ggplot() +
    aes(x = amount.ug, y = depth) +
    geom_point() +
    #facet_wrap(~peak_name, scales = "free") +
    scale_y_reverse() 

```

###Isorenieretane
```{r "Isorenieretane_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='Isorenieretane') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```


###Chlorobactane
```{r "chlorobactane_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='Chlorobactane') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```

###all Aryls
```{r "aryls_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='all_Aryl_Isop') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```


###all PAHs
```{r "overview_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='all_PAH') %>%
  #filter(!is.na(depth==108)) %>%
    ggplot() +
    aes(x = conc_tle.ug.g, y = depth) +
    geom_point() +
    #facet_wrap(~peak_name, scales = "free") 
    scale_y_reverse() 

```

###3 ring PAHs
```{r "overview_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='3ring_all') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```
###4 ring PAHs
```{r "overview_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='4ring_all') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```

###5 ring PAHs
```{r "overview_against_depth", fig.width = 10, fig.height = 8}

subset(final_data, peak_name=='5ring_all') %>%
  ggplot() +
  aes(x = conc_tle.ug.g, y = depth) +
  geom_point() +
  #facet_wrap(~peak_name, scales = "free") 
  scale_y_reverse() 

```

